---
title: "Population change map"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interactive-map}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(testmap)
```


```{r load-packages, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(leaflet)
library(leafem)
library(sf)
```

Use the interactive map below to check population changes. Zoom in to see 
where *Podocnemis unilfilis* is Endangered based on IUCN Redlist criteria - A3bd.

This is an extension of [Norris et al 2019](https://doi.org/10.1016/j.biocon.2019.02.022) that includes:

- stochastic population projections 
- threats caused by human acessibility (hunting and habitat change) and actions that reduce river connectivity.

When you zoom in you will see brown shaded points that show where 
populations are expected to decline by 50% or more in 42 years
```{r make-map, echo=FALSE, message=FALSE, warning=FALSE}
points_bau4326 <- points_bau |> 
  #filter(BASIN_N == "Orinoco") |>
  arrange(BASIN_N, SUBBASI, BB_ID, REACH_ID) |> 
  filter(row_number() %% 5 == 1) |> st_as_sf() |> st_transform(4326) 
levels(points_bau4326$flag_50_42y) <- c("No", "Yes")
points_bau4326_low <- points_bau4326
# colour palette
leaf_pal <- colorFactor(
  palette = c("#7274C1", "#A3720E"), 
  domain = points_bau4326$flag_50_42y
)

mypal <- c("#7274C1", "#A3720E")
 # basemap with tile control. not working.
basemap <- leaflet::leaflet() %>%
  # add different provider tiles
  addProviderTiles(
    "OpenStreetMap",
    # give the layer a name
    group = "OpenStreetMap"
  ) %>%
  addProviderTiles(
    "Stamen.Terrain",
    group = "Stamen.Terrain"
  ) %>%
  addProviderTiles(
    "Esri.WorldImagery",
    group = "Esri.WorldImagery"
  ) %>%
# add a layers control
  addLayersControl(
    baseGroups = c(
      "OpenStreetMap", "Stamen.Terrain", "Esri.WorldImagery"
    ),
    # position it on the topleft
    position = "topleft"
  )

# interactive map
leaflet::leaflet(points_bau4326_low) |> 
  addTiles() |> 
  addCircles(color = ~leaf_pal(flag_50_42y), 
             group = "points_bau4326_low"
) |>  
  addCircleMarkers(color = ~leaf_pal(flag_50_42y), 
             stroke = FALSE, fillOpacity = 0.4, 
clusterOptions = markerClusterOptions(), 
group = "points_bau4326"
) |> 
  addLegend("bottomright", pal = leaf_pal, title="Endangered",
            values = ~flag_50_42y,
            group = "en_legend") |> 
  groupOptions("points_bau4326", zoomLevels = 1:8) |> 
  groupOptions("points_bau4326_low", zoomLevels = 9:15) |> 
  groupOptions("en_legend", zoomLevels = 9:15) |> 
  addScaleBar() |> 
  leafem::addMouseCoordinates()
  
  
```

