---
title: "Population change map"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interactive-map}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!-- Google Translate widget -->
<div id="google_translate_element" style="float: right;"></div>
<script type="text/javascript">
function googleTranslateElementInit() {
  new google.translate.TranslateElement({pageLanguage: 'en', includedLanguages: 'ar,bn,pt,en,nl,es,fr,de,it,iw,hi,ja,ko,zh-CN'}, 'google_translate_element');
}
</script>
<script type="text/javascript" src="//translate.google.com/translate_a/element.js?cb=googleTranslateElementInit"></script>


```{r setup, echo=FALSE}
library(testmap)
```


```{r load-packages, echo=FALSE, message=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(ggplot2)
library(leaflet)
library(leafem)
library(Hmisc)
library(sf)
library(knitr)
library(kableExtra)
library(htmltools) 
library(patchwork)
```

```{r main-result, echo=FALSE}
tot_t0 <- sum(points_bau_ffr$fem_t0)
tot_t35 <- floor(sum(points_bau_ffr$fem_t35))
pop_change_35 <- (tot_t35 - tot_t0) / tot_t0 
change_value <- round(abs(pop_change_35) * 100, 1)

# quantile wont work as there are few distinct values.
# hist(points_bau_ffr$fem_diff_t35)
#pop_change_35_q25 <- quantile(points_bau_ffr$fem_diff_t35, probs =0.25)
#pop_change_35_q75 <- quantile(points_bau_ffr$fem_diff_t35, probs =0.75)

# 100 years
tot_t100 <- floor(sum(points_bau_ffr$fem_t100))
pop_change_100 <- (tot_t100 - tot_t0) / tot_t0 
change_value_100 <- round(abs(pop_change_100) * 100, 1)


# Randomized resample to get quantile distribution.
# Sample Function
resample_and_calculate_change <- function(df, sample_percent = NA, 
                                          sample_size = NA, 
                                          col1, col2) {
  # sample
  if(!is.na(sample_size)){ 
    sampled_df <- df |> 
    dplyr::slice_sample(n = sample_size, replace = TRUE) 
    }else{
  sampled_df <- df |> 
    dplyr::slice_sample(prop = (sample_percent / 100)) 
    }
  # calculate change
  tot_tstart <- sum(sampled_df[ , col1]) 
  tot_tend <- floor(sum(sampled_df[ , col2])) 
  pop_change <- (tot_tend - tot_tstart) / tot_tstart
  return(pop_change)
}

# 3 generations - 35 year is shortest (25 percentile of matrix model mother to cohort ages).
set.seed(123)  # For reproducibility
results_35 <- replicate(10000, resample_and_calculate_change(df = points_bau_ffr, sample_size = 1000, 
                                                      col1 = "fem_t0", col2 = "fem_t35")) 
pop_change_35_q25 <- quantile(results_35, probs =0.25)
pop_change_35_q75 <- quantile(results_35, probs =0.75)
q25_t35 <- round(abs(pop_change_35_q25) * 100, 1)
q75_t35 <- round(abs(pop_change_35_q75) * 100, 1)

# 100 years
results_100 <- replicate(10000, resample_and_calculate_change(df = points_bau_ffr, 
                                                              sample_size = 1000, 
                                                      col1 = "fem_t0", col2 = "fem_t100")) 
pop_change_100_q25 <- quantile(results_100, probs =0.25)
pop_change_100_q75 <- quantile(results_100, probs =0.75)
q25_t100 <- round(abs(pop_change_100_q25) * 100, 1)
q75_t100 <- round(abs(pop_change_100_q75) * 100, 1)

```

Use the interactive map below to check population changes. Zoom in to see 
where *Podocnemis unifilis* is Endangered based on IUCN Red List criteria - A3bd.

- Overall *Podocnemis unifilis* is Endangered based on future population 
size reduction criteria.    
Within 3 generations (35 years) the adult female population is predicted 
to decline by `r change_value`% in the future (25 and 75% quantile 
range: `r q25_t35` - `r q75_t35` % decline). Within 100 years the adult 
female population is predicted to decline by `r change_value_100`%.

## Map
Due to the number of points, the map can become slow to respond when you zoom in.
Zoom out to a level showing fewer points and you can pan around the map 
to find the area of interest. Then zoom in again to check the coloured points.

If you find any errors (e.g. points where *Podocnemis unifilis* does not occur) 
please raise an issue at: https://github.com/darrennorris/testmap/issues . 

```{r make-map-ffr, echo=FALSE, message=FALSE, warning=FALSE}
points_bau4326 <- sf::st_as_sf(points_bau_ffr_map, crs = 3395) |> 
  st_transform(4326) 
levels(points_bau4326$flag_50_35y) <- c("No", "Yes")
points_bau4326_low <- points_bau4326
# label to plot with circle
points_bau4326_low$label <- paste("Prot = ", points_bau4326_low$Protected, 
                                  "Acc = ", points_bau4326_low$Accessible, 
                                  "Free = ", points_bau4326_low$Free_flowing)
# colour palette
leaf_pal <- colorFactor(
  palette = c("#7274C1", "#A3720E"), 
  domain = points_bau4326$flag_50_35y
)

mypal <- c("#7274C1", "#A3720E")

# interactive map. Options added to make panning smoother....
leaflet::leaflet(points_bau4326_low, 
                 options = leafletOptions(preferCanvas = TRUE)) |> 
  addTiles(options = tileOptions(
  updateWhenZooming = FALSE,      # map won't update tiles until zoom is done
  updateWhenIdle = TRUE)) |> 
  addCircles(color = ~leaf_pal(flag_50_35y), 
             popup = ~htmlEscape(label),
             group = "points_bau4326_low"
) |>  
  addCircleMarkers(color = ~leaf_pal(flag_50_35y), 
             stroke = FALSE, fillOpacity = 0.4, 
clusterOptions = markerClusterOptions(), 
group = "points_bau4326"
) |> 
  addLegend("bottomright", pal = leaf_pal, title="Endangered",
            values = ~flag_50_35y,
            group = "en_legend") |> 
  groupOptions("points_bau4326", zoomLevels = 1:7) |> 
  groupOptions("points_bau4326_low", zoomLevels = 8:15) |> 
  groupOptions("en_legend", zoomLevels = 8:15) |> 
  addScaleBar() |> 
  leafem::addMouseCoordinates()
  
```

When you zoom in you will see shaded points. The points follow rivers produced by
remote sensing (Free-flowing Rivers, 2019: https://doi.org/10.1038/s41586-019-1111-9). 
To facilitate online viewing the mapped points are a subset at intervals 
of approximately 10 kilometers.
The points are brown where populations are predicted to decline by 50% 
or more within 3 generations (35 years). Brown points therefore represent 
rivers where the species is Endangered, following the IUCN Red List 
population size reduction criteria - A3bd (https://www.iucnredlist.org/about/faqs).

The points represent rivers where *Podocnemis unifilis* females are 
likely to nest and that are likely to be accessible to people by boat. 

```{r make-map-norris, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
# NOT USED.
# Map below uses the river points from Norris et. al. 2019.
points_bau4326 <- points_bau |> 
  #filter(BASIN_N == "Orinoco") |>
  arrange(BASIN_N, SUBBASI, BB_ID, REACH_ID) |> 
  filter(row_number() %% 5 == 1) |> st_as_sf(crs = 3395) |> st_transform(4326) 
levels(points_bau4326$flag_50_35y) <- c("No", "Yes")
points_bau4326_low <- points_bau4326
# label to plot with circle
points_bau4326_low$label <- paste("Prot = ", points_bau4326_low$Protected, 
                                  "Acc = ", points_bau4326_low$Accessible, 
                                  "Free = ", points_bau4326_low$Free_flowing)
# colour palette
leaf_pal <- colorFactor(
  palette = c("#7274C1", "#A3720E"), 
  domain = points_bau4326$flag_50_35y
)

mypal <- c("#7274C1", "#A3720E")
 # basemap with tile control. not working.
basemap <- leaflet::leaflet() %>%
  # add different provider tiles
  addProviderTiles(
    "OpenStreetMap",
    # give the layer a name
    group = "OpenStreetMap"
  ) %>%
  addProviderTiles(
    "Stamen.Terrain",
    group = "Stamen.Terrain"
  ) %>%
  addProviderTiles(
    "Esri.WorldImagery",
    group = "Esri.WorldImagery"
  ) %>%
# add a layers control
  addLayersControl(
    baseGroups = c(
      "OpenStreetMap", "Stamen.Terrain", "Esri.WorldImagery"
    ),
    # position it on the topleft
    position = "topleft"
  )

# interactive map. Optins added to make panning smoother....
leaflet::leaflet(points_bau4326_low, 
                 options = leafletOptions(preferCanvas = TRUE)) |> 
  addTiles(options = tileOptions(
  updateWhenZooming = FALSE,      # map won't update tiles until zoom is done
  updateWhenIdle = TRUE)) |> 
  addCircles(color = ~leaf_pal(flag_50_35y), 
             popup = ~htmlEscape(label),
             group = "points_bau4326_low"
) |>  
  addCircleMarkers(color = ~leaf_pal(flag_50_35y), 
             stroke = FALSE, fillOpacity = 0.4, 
clusterOptions = markerClusterOptions(), 
group = "points_bau4326"
) |> 
  addLegend("bottomright", pal = leaf_pal, title="Endangered",
            values = ~flag_50_35y,
            group = "en_legend") |> 
  groupOptions("points_bau4326", zoomLevels = 1:8) |> 
  groupOptions("points_bau4326_low", zoomLevels = 9:15) |> 
  groupOptions("en_legend", zoomLevels = 9:15) |> 
  addScaleBar() |> 
  leafem::addMouseCoordinates()
  
  
```


## Methods
The analysis is developed here: https://github.com/darrennorris/TACAR, 
The methods used are an extension of [Norris et. al. 2019](https://doi.org/10.1016/j.biocon.2019.02.022) that includes:

- Stochastic population projections. 
- Future impacts to populations caused by human acessibility (hunting) and actions that reduce river connectivity.


## Summaries
A summary of the map results is presented below. These include summaries 
by country and major basin using the full set of points at 1 kilometer 
intervals. To date subpopulations have not been identified due to a lack of 
comprehensive genetic analysis. Summaries by basin are provided as these 
are expected to represent important management units for this predominantly 
aquatic species. The implementation of conservation actions is regulated 
at national levels so summaries are also provided by country.


 - Results include the size and extent of future population changes.    
 The column "population change" is the difference in
the number of adult females after 3 generations (35 years). This is consistent 
with the population size reduction criteria used by the IUCN Red List. The 
column "River length Endangered", is the proportion of river length where 
populations declined by 50% or more, and represents the spatial extent of 
future changes. 

```{r totals-access, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
# Check length of free flowing by Accessibility.
points_bau |> 
  mutate(flag_ff = if_else(Free_flowing == "yes", 1, 0)) |>
  group_by(Accessible) |> 
  summarise(length_river = n(), 
           length_ff = sum(flag_ff)) |> 
  ungroup() |> 
  mutate(prop_ff = round((length_ff / length_river), 2), 
         length_notff = (length_river - length_ff)) |> 
  select(Accessible, prop_ff, length_river, length_ff, length_notff) 
  
```

### Summary by country.

```{r totals-country, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
# Not used. Updated points from Norris et al. 2019.
table_country <- points_bau |> 
  mutate(flag_ff = if_else(Free_flowing == "yes", 1, 0)) |>
  group_by(COUNTRY, Accessible) |> 
  mutate(flag_e = if_else(fem_diff_t35 <= -0.5, 1, 0)) |>
  summarise(length_river = n(), 
            length_end = sum((flag_e)), 
           length_ff = sum(flag_ff), 
           prop_ff = round((sum(flag_ff) / length_river), 2)) |> 
    ungroup() |> 
    mutate(length_not_end = length_river - length_end, 
           prop_end = round((length_end / length_river), 2), 
           length_not_ff = length_river - length_ff
           
) |> 
  select(COUNTRY, Accessible, prop_end, length_river, length_end, length_not_end, 
         prop_ff, length_ff, length_not_ff)

# Sum rows of each column if numeric 
func <- function(z) if (is.numeric(z)) sum(z) else '' 
sumrow <- as.data.frame(lapply(table_country, func))
# Give name to the first element of the new data frame created above
sumrow[1] <- "Total"
# remove unwanted value
sumrow[c(2, 3, 7)] <- NA
```

```{r sample-wrapper, echo=FALSE, warning=FALSE, message=FALSE}
# wrapper to so can use sample function with plyr::ddply
wrap_sample <- function(x){ 
results <- replicate(1000, resample_and_calculate_change(df = x, sample_size = 1000, 
                                                      col1 = "fem_t0", col2 = "fem_t35")) 
pop_change_35_q25 <- quantile(results, probs =0.25)
pop_change_35_q75 <- quantile(results, probs =0.75)

dfout <- data.frame(q25_value = round(pop_change_35_q25, 3), 
           q75_value = round(pop_change_35_q75, 3))
return(dfout)
  }
```


```{r totals-country-ffr, echo=FALSE, warning=FALSE, message=FALSE}
# All FFR "yes". Need to revise and update.
# interquartile range
df_quartile <- plyr::ddply(points_bau_ffr, c("COUNTRY"), .fun = wrap_sample)
# other summaries
table_country_ffr <- points_bau_ffr |> 
  mutate(flag_ff = if_else(Free_flowing == "yes", 1, 0)) |>
  group_by(COUNTRY) |> 
  mutate(flag_e = if_else(fem_diff_t35 <= -0.5, 1, 0)) |>
  summarise(pop_change = round(((sum(fem_t35) - sum(fem_t0)) / sum(fem_t0)), 2), 
            #change_lcl_95 = Hmisc::smean.cl.boot(fem_diff_t35)["Lower"], 
            #change_ucl_95 = Hmisc::smean.cl.boot(fem_diff_t35)["Upper"],
            length_river = n(), 
            length_end = sum((flag_e)), 
           length_ff = sum(flag_ff), 
           prop_ff = round((sum(flag_ff) / length_river), 2)) |> 
    ungroup() |> 
  left_join(df_quartile) |>
    mutate(length_not_end = length_river - length_end, 
           prop_end = round((length_end / length_river), 2), 
           length_not_ff = length_river - length_ff, 
           change_label = paste(pop_change, " (", round(q25_value, 2), " ", 
                                "\u2013", " ", round(q75_value, 2), ")", sep="")
           
) |> 
  select(COUNTRY, change_label, prop_end, length_river, length_end, length_not_end, 
         prop_ff, length_ff, length_not_ff)

# Sum rows of each column if numeric 
func <- function(z) if (is.numeric(z)) sum(z) else '' 
sumrow_ffr <- as.data.frame(lapply(table_country_ffr, func))
# Give name to the first element of the new data frame created above
sumrow_ffr[1] <- "Total"
# remove unwanted value
sumrow_ffr[c(3, 7)] <- NA

```


```{r make-table-country-ffr, echo=FALSE}
options(knitr.kable.NA = '')

bind_rows(table_country_ffr, sumrow_ffr) |> 
  select(COUNTRY, change_label, prop_end, length_river) |>
knitr::kable(col.names = c("Country", "Population change", 
                           "River length Endangered (prop)", 
                           "River length (km)"), 
             caption = "Table 1. Population changes by country."
             )|> 
  kableExtra::column_spec(3:4, width = "3.0cm")
             
```



### Summary by basin.
In the table below "Amazon" includes the Amazon Basin together 
with the Araguaia-Tocantins Basin in Brazil. "Coastal North" covers 
the Guianas and includes the Essequibo Basin. "Coastal South" includes 
part of Maranhão State in Brazil.
```{r totals-basin, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
# Not used. 
# Results from updated Norris et. al. 2019.
table_basin <- points_bau |> 
  group_by(BASIN_N) |> 
  mutate(flag_e = if_else(fem_diff_t35 <= -0.5, 1, 0)) |>
  summarise(pop_change = round(((sum(fem_t35) - sum(fem_t0)) / sum(fem_t0)), 3), 
            length_river = n(), 
            length_end = sum((flag_e))) |> 
    ungroup() |> 
    mutate(length_not_end = length_river - length_end, 
           prop_end = round((length_end / length_river), 2)) |> 
  select(BASIN_N, pop_change, prop_end, length_river, length_end, length_not_end)
```

```{r make-table-basin, echo=FALSE, eval=FALSE}
# not used
knitr::kable(table_basin, 
             col.names = c("Basin", "Population change", "Endangered (prop)","river length (km)", 
                           "length End.", "length not End.")) 
             
```


```{r totals-basin-ffr, echo=FALSE, message=FALSE, warning=FALSE}
# interquartile range
df_quartile_b <- plyr::ddply(points_bau_ffr, c("BASIN_NAME"), .fun = wrap_sample)

table_basin_ffr <- points_bau_ffr |> 
  group_by(BASIN_NAME) |> 
  mutate(flag_e = if_else(fem_diff_t35 <= -0.5, 1, 0)) |>
  summarise(pop_start = sum(fem_t0), 
            pop_end = sum(fem_t35), 
            pop_change = round(((sum(fem_t35) - sum(fem_t0)) / sum(fem_t0)), 3),
            #change_lcl_95 = Hmisc::smean.cl.boot(fem_diff_t35)["Lower"], 
            #change_ucl_95 = Hmisc::smean.cl.boot(fem_diff_t35)["Upper"],
            length_river = n(), 
            length_end = sum((flag_e))) |> 
    ungroup() |> 
  left_join(df_quartile_b) |>
    mutate(length_not_end = length_river - length_end, 
           prop_end = round((length_end / length_river), 2), 
           change_label = paste(pop_change, " (", round(q25_value, 2), " ", 
                                "\u2013", " ", round(q75_value, 2), ")", sep="")
           ) |> 
  select(BASIN_NAME, change_label, prop_end, length_river, length_end, length_not_end)
```

```{r make-table-basin-ffr, echo=FALSE}
table_basin_ffr |> 
  select(BASIN_NAME, change_label, prop_end, length_river) |>
knitr::kable(col.names = c("Basin", "population change", 
                           "River length Endangered (prop)", 
                           "River length (km)"), 
             caption = "Table 2. Population changes by basin.") 
             
```


### Summary by major basins and countries.
```{r totals-basin-country, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE, }
# Not used.
table_basin_country <- points_bau |> 
  group_by(BASIN_N, COUNTRY) |> 
  mutate(flag_e = if_else(fem_diff_t35 <= -0.5, 1, 0)) |>
  summarise(pop_change = round(((sum(fem_t35) - sum(fem_t0)) / sum(fem_t0)), 3), 
            length_river = n(), 
            length_end = sum((flag_e))) |> 
    ungroup() |> 
    mutate(length_not_end = length_river - length_end, 
           prop_end = round((length_end / length_river), 2)) |> 
  select(BASIN_N, COUNTRY, pop_change, prop_end, length_river, length_end, length_not_end) 

```

```{r make-table-basin-country, echo=FALSE, eval=FALSE}
# Not used.
knitr::kable(table_basin_country, 
             col.names = c("Basin", "Country", 
                           "Population change","Endangered (prop)", 
                           "river length (km)", 
                           "length End.", "length not End."), 
             caption = "Table 3. Population changes by country and basin.") |> 
  kableExtra::column_spec(3:7, width = "3cm")
             
```


```{r data-basin-country-ffr, echo=FALSE, message=FALSE, warning=FALSE}
# Make summaries
# interquartile range
df_quartile_cb <- plyr::ddply(points_bau_ffr, c("COUNTRY", "BASIN_NAME"), .fun = wrap_sample)
# other summaries
tab_sum <- points_bau_ffr |> 
  dplyr::mutate(flag_EN = if_else(fem_diff_t35 <= -0.5, 1, 0)) |>
  dplyr::group_by(COUNTRY, BASIN_NAME) |> 
  dplyr::summarise(pop_start = sum(fem_t0), 
                   pop_end = sum(fem_t35), 
                   pop_change = round(((sum(fem_t35) - sum(fem_t0)) / sum(fem_t0)), 2),
                   diff_mean = mean(fem_diff_t35), 
                   #change_lcl_95 = Hmisc::smean.cl.boot(fem_diff_t35)["Lower"], 
                   #change_ucl_95 = Hmisc::smean.cl.boot(fem_diff_t35)["Upper"],
                   diff_median = median(fem_diff_t35),
                   diff_q25 = quantile(fem_diff_t35, probs = 0.25), 
                   diff_q75 = quantile(fem_diff_t35, probs = 0.75),
                   length_river = n(), 
                   length_endangered = sum((flag_EN))) |>
  dplyr::ungroup() |> 
  left_join(df_quartile_cb) |>
  dplyr::mutate(proportion_endangered = round((length_endangered / length_river), 2), 
                length_label = round((length_river / 1000), 1)
                ) |> 
  dplyr::mutate(threat_status = case_when(diff_mean <= -0.8 ~ "Critically Endangered", 
                                   diff_mean <= -0.5 ~ "Endangered", 
                                   diff_mean <= -0.3 ~ "Vulnerable", 
                                   diff_mean <= -0.2 ~ "Near Threatened"), 
                change_label = paste(pop_change, " (", round(q25_value, 2), " ", 
                                "\u2013", " ", round(q75_value, 2), ")", sep="")
           ) 
#tab_sum$bc <- factor(paste(tab_sum$BASIN_NAME, tab_sum$COUNTRY, sep = "\n"))
# levels(tab_sum$bc)

river_tot <- tab_sum |> 
  filter(length_river > 21) |> 
  droplevels() |> 
  pull(length_river) |> sum() 
river_amazon_brazil <-   tab_sum |>
  filter(COUNTRY == "Brazil", BASIN_NAME == "Amazon") |> 
  droplevels() |> 
  pull(length_river) |> sum() 
river_amazon_brazil_prop <- round(((river_amazon_brazil / river_tot) * 100), 1)
```


```{r make-table-basin-country-ffr, echo=FALSE}
tab_sum |> 
  filter(length_river > 21) |> 
  droplevels() |>
  select(BASIN_NAME, COUNTRY, change_label, proportion_endangered, 
         length_river) |> 
  arrange(BASIN_NAME, COUNTRY) |>
knitr::kable(col.names = c("Basin", "Country", "population change", 
                           "River length Endangered (prop)", 
                           "river length (km)"), 
             caption = "Table 3. Population changes by country and basin.") |> 
  kableExtra::column_spec(4:5, width = "3cm")
             
```


The table values can also be presented graphically.
This enables a clearer understanding of the patterns.

```{r fig-basin-country-ffr, fig.cap = "Figure 1. Population changes across basins and contries.", echo=FALSE, message=FALSE, warning=FALSE}
# Plot
# Add width ajustment factor and fill colour to be consistent across graphs.
mycountry <- sort(unique(tab_sum$COUNTRY))
ncountry <- length(unique(tab_sum$COUNTRY))
palette_country <- grey.colors(n = ncountry)
tab_bc <- tab_sum |> 
  filter(length_river > 21) |> 
  droplevels() |>
  left_join(data.frame(COUNTRY = mycountry, 
                       my_fill = palette_country))
# Now include width factor to be conistent across facets
tab_bc <- tab_bc |> left_join(tab_bc |> 
  group_by(BASIN_NAME) |> summarise(country_count = n()) |> 
  ungroup() |> mutate(width_fac = country_count / max(country_count)) 
)
tab_bc$river_log <- log1p((tab_bc$length_river/1000))
tab_bc$river_fac <- tab_bc$river_log / max(tab_bc$river_log)
# check where these errors come from - 
# Orinoco in Brazil, Orinoco in Guyana, Coastal South in French Guiana. 
fig_left <- tab_bc |> 
  filter(BASIN_NAME %in% c("Amazon", "Coastal North")) |>
ggplot(aes(x = COUNTRY, y = pop_change, width = (.85 * river_fac)*width_fac)) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.2, ymax = -0.3,
            fill="#F6DABF", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.3, ymax = -0.5,
            fill="#DEB688", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.5, ymax = -0.8,
            fill="#C09355", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.8, ymax = -1,
            fill="#A3720E", geom = 'rect', alpha=0.4) + 
  geom_hline(yintercept = 0, colour = "black") +
  geom_col(aes(fill = my_fill)) + 
  geom_errorbar(aes(ymax = q75_value, ymin = q25_value), 
                width = 0.2) + 
  geom_label(aes(x = COUNTRY, y = 0.2, label = length_label)) +
  scale_fill_identity() +  
  facet_wrap(~BASIN_NAME, scales = "free_x", nrow = 2) +
  theme_bw() + 
  labs(y = "Population change", x = "") + 
  #theme(legend.position = "bottom")
  guides(fill = "none")

fig_right <- tab_bc |> 
  filter(BASIN_NAME %in% c("Coastal South", "Orinoco")) |>
ggplot(aes(x = COUNTRY, y = pop_change, width = (.85 * river_fac)*(width_fac*3))) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.2, ymax = -0.3,
            fill="#F6DABF", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.3, ymax = -0.5,
            fill="#DEB688", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.5, ymax = -0.8,
            fill="#C09355", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.8, ymax = -1,
            fill="#A3720E", geom = 'rect', alpha=0.4) + 
  geom_hline(yintercept = 0, colour = "black") +
  geom_col(aes(fill = my_fill)) + 
  geom_errorbar(aes(ymax = q75_value, ymin = q25_value), 
                width = 0.2) + 
  geom_label(aes(x = COUNTRY, y = 0.2, label = length_label)) + 
  scale_fill_identity() + 
  facet_wrap(~BASIN_NAME, scales = "free_x", nrow = 2) +
  theme_bw() + 
  labs(y = "", x = "") + 
  #theme(legend.position = "bottom")
  guides(fill = "none") 
# put together
fig_left + fig_right + plot_layout(widths = c(3, 1))

```

The grey bars in the above figure show predicted population changes for each basin country combination.
The width of the bars is proportional to the river length on a log scale. 
The labels above each bar show river lengths in thousands of 
kilometers (e.g. "22.2" is 22,200 kilometers).
The background shading follows a gradient from Near Threatened ("-0.2" is 20% loss) to 
Critically Endangered ("-0.8" is 80% loss). 

The Amazon Basin in Brazil represents the largest proportion of 
rivers (`r river_amazon_brazil_prop`%). Considering the spatial variation 
in threats and genetics, it would probably be useful to 
include the Amazon Basin seperated into major basins (Madeira, Negro, etc). 
But the summaries presented here provide a general overview and enable 
assessment of the results, and the overall patterns do not change if the 
Amazon basin is seperated further.

## Next steps

 - Consult with researchers to check and validate the mapped results.
 - Evaluate future impact of conservation actions.
 - Apply to other species.
