---
title: "Population change map"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interactive-map}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!-- Google Translate widget -->
<div id="google_translate_element" style="float: right;"></div>
<script type="text/javascript">
function googleTranslateElementInit() {
  new google.translate.TranslateElement({pageLanguage: 'en', includedLanguages: 'ar,bn,pt,en,nl,es,fr,de,it,iw,hi,ja,ko,zh-CN'}, 'google_translate_element');
}
</script>
<script type="text/javascript" src="//translate.google.com/translate_a/element.js?cb=googleTranslateElementInit"></script>


```{r setup, echo=FALSE}
library(testmap)
```


```{r load-packages, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(leaflet)
library(leafem)
library(Hmisc)
library(sf)
library(knitr)
library(kableExtra)
library(htmltools) 
library(patchwork)
```

```{r main-result, echo=FALSE}
tot_t0 <- sum(points_bau_ffr$fem_t0)
tot_t42 <- floor(sum(points_bau_ffr$fem_t42))
pop_change_42 <- (tot_t42 - tot_t0) / tot_t0 
change_value <- round(abs(pop_change_42) * 100, 1)
diff_boot <- Hmisc::smean.cl.boot(points_bau_ffr$fem_diff_t42)
pop_change_42_lcl <- diff_boot["Lower"]
pop_change_42_ucl <- diff_boot["Upper"]
# quantile wont work as there are few distinct values.
# hist(points_bau_ffr$fem_diff_t42)
#pop_change_42_q25 <- quantile(points_bau_ffr$fem_diff_t42, probs =0.25)
#pop_change_42_q75 <- quantile(points_bau_ffr$fem_diff_t42, probs =0.75)

# Randomized resample to get quantile distribution.

# Sample Function
resample_and_calculate_change <- function(df, sample_percent, col1, col2) {
  # sample
  sampled_df <- points_bau_ffr |> 
    dplyr::slice_sample(prop = (sample_percent / 100))
  # calculate change
  tot_tstart <- sum(sampled_df[ , col1]) 
  tot_tend <- floor(sum(sampled_df[ , col2])) 
  pop_change <- (tot_tend - tot_tstart) / tot_tstart
  return(pop_change)
}

set.seed(123)  # For reproducibility
results <- replicate(10000, resample_and_calculate_change(points_bau_ffr, 1, 
                                                       "fem_t0", "fem_t42")) 
pop_change_42_q25 <- quantile(results, probs =0.25)
pop_change_42_q75 <- quantile(results, probs =0.75)
q25_value <- round(abs(pop_change_42_q25) * 100, 1)
q75_value <- round(abs(pop_change_42_q75) * 100, 1)
```

Use the interactive map below to check population changes. Zoom in to see 
where *Podocnemis unilfilis* is Endangered based on IUCN Red List criteria - A3bd.

- Overall *Podocnemis unilfilis* is Endangered based on future population 
size reduction criteria. 
The adult female population is predicted to decline by `r change_value`%
in the future (25 and 75% quantile range: `r q25_value` - `r q75_value` % decline).

The analysis is developed here: https://github.com/darrennorris/TACAR, 
and is an extension of [Norris et. al. 2019](https://doi.org/10.1016/j.biocon.2019.02.022) that includes:

- Stochastic population projections. 
- Future impacts to populations caused by human acessibility (hunting) and actions that reduce river connectivity.


## Map
Due to the number of points, the map can become slow to respond when you zoom in.
Zoom out to a level showing fewer points and you can pan around the map 
to find the area of interest. Then zoom in again to check the coloured points.

```{r make-map-ffr, echo=FALSE, message=FALSE, warning=FALSE}
points_bau4326 <- sf::st_as_sf(points_bau_ffr_map, crs = 3395) |> 
  st_transform(4326) 
levels(points_bau4326$flag_50_42y) <- c("No", "Yes")
points_bau4326_low <- points_bau4326
# label to plot with circle
points_bau4326_low$label <- paste("Prot = ", points_bau4326_low$Protected, 
                                  "Acc = ", points_bau4326_low$Accessible, 
                                  "Free = ", points_bau4326_low$Free.flowing)
# colour palette
leaf_pal <- colorFactor(
  palette = c("#7274C1", "#A3720E"), 
  domain = points_bau4326$flag_50_42y
)

mypal <- c("#7274C1", "#A3720E")

# interactive map. Options added to make panning smoother....
leaflet::leaflet(points_bau4326_low, 
                 options = leafletOptions(preferCanvas = TRUE)) |> 
  addTiles(options = tileOptions(
  updateWhenZooming = FALSE,      # map won't update tiles until zoom is done
  updateWhenIdle = TRUE)) |> 
  addCircles(color = ~leaf_pal(flag_50_42y), 
             popup = ~htmlEscape(label),
             group = "points_bau4326_low"
) |>  
  addCircleMarkers(color = ~leaf_pal(flag_50_42y), 
             stroke = FALSE, fillOpacity = 0.4, 
clusterOptions = markerClusterOptions(), 
group = "points_bau4326"
) |> 
  addLegend("bottomright", pal = leaf_pal, title="Endangered",
            values = ~flag_50_42y,
            group = "en_legend") |> 
  groupOptions("points_bau4326", zoomLevels = 1:7) |> 
  groupOptions("points_bau4326_low", zoomLevels = 8:15) |> 
  groupOptions("en_legend", zoomLevels = 8:15) |> 
  addScaleBar() |> 
  leafem::addMouseCoordinates()
  
```

When you zoom in you will see shaded points. The points follow rivers produced by
remote sensing (Free-flowing Rivers, 2019: https://doi.org/10.1038/s41586-019-1111-9). 
To facilitate online viewing the mapped points are a subset at intervals 
of approximately 10 kilometers.
The points are brown where populations are predicted to decline by 50% 
or more within 3 generations (42 years). Brown points therefore represent 
rivers where the species is Endangered, following the IUCN Red List 
population size reduction criteria - A3bd (https://www.iucnredlist.org/about/faqs).

```{r make-map-norris, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
#Map below uses the river points from Norris et. al. 2019.
points_bau4326 <- points_bau |> 
  #filter(BASIN_N == "Orinoco") |>
  arrange(BASIN_N, SUBBASI, BB_ID, REACH_ID) |> 
  filter(row_number() %% 5 == 1) |> st_as_sf(crs = 3395) |> st_transform(4326) 
levels(points_bau4326$flag_50_42y) <- c("No", "Yes")
points_bau4326_low <- points_bau4326
# label to plot with circle
points_bau4326_low$label <- paste("Prot = ", points_bau4326_low$Protected, 
                                  "Acc = ", points_bau4326_low$Accessible, 
                                  "Free = ", points_bau4326_low$Free.flowing)
# colour palette
leaf_pal <- colorFactor(
  palette = c("#7274C1", "#A3720E"), 
  domain = points_bau4326$flag_50_42y
)

mypal <- c("#7274C1", "#A3720E")
 # basemap with tile control. not working.
basemap <- leaflet::leaflet() %>%
  # add different provider tiles
  addProviderTiles(
    "OpenStreetMap",
    # give the layer a name
    group = "OpenStreetMap"
  ) %>%
  addProviderTiles(
    "Stamen.Terrain",
    group = "Stamen.Terrain"
  ) %>%
  addProviderTiles(
    "Esri.WorldImagery",
    group = "Esri.WorldImagery"
  ) %>%
# add a layers control
  addLayersControl(
    baseGroups = c(
      "OpenStreetMap", "Stamen.Terrain", "Esri.WorldImagery"
    ),
    # position it on the topleft
    position = "topleft"
  )

# interactive map. Optins added to make panning smoother....
leaflet::leaflet(points_bau4326_low, 
                 options = leafletOptions(preferCanvas = TRUE)) |> 
  addTiles(options = tileOptions(
  updateWhenZooming = FALSE,      # map won't update tiles until zoom is done
  updateWhenIdle = TRUE)) |> 
  addCircles(color = ~leaf_pal(flag_50_42y), 
             popup = ~htmlEscape(label),
             group = "points_bau4326_low"
) |>  
  addCircleMarkers(color = ~leaf_pal(flag_50_42y), 
             stroke = FALSE, fillOpacity = 0.4, 
clusterOptions = markerClusterOptions(), 
group = "points_bau4326"
) |> 
  addLegend("bottomright", pal = leaf_pal, title="Endangered",
            values = ~flag_50_42y,
            group = "en_legend") |> 
  groupOptions("points_bau4326", zoomLevels = 1:8) |> 
  groupOptions("points_bau4326_low", zoomLevels = 9:15) |> 
  groupOptions("en_legend", zoomLevels = 9:15) |> 
  addScaleBar() |> 
  leafem::addMouseCoordinates()
  
  
```


## Summaries
A summary of the map results is presented below. This includes summaries 
by country and major basin using the full set of points at 1 kilometer 
intervals. Results include the size and extent of likely
population changes. The column "population change" is the difference in
the number of adult females after 3 generations (42 years). This is consistent 
with the population sze reduction criteria used by the IUCN Red List. The 
column "River length Endangered", is the proportion of river length where 
populations declined by 50% or more, and represents the spatial extent of 
future changes. 

```{r totals-access, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
# Check length of free flowing by Accessibility.
points_bau |> 
  mutate(flag_ff = if_else(Free.flowing == "yes", 1, 0)) |>
  group_by(Accessible) |> 
  summarise(length_river = n(), 
           length_ff = sum(flag_ff)) |> 
  ungroup() |> 
  mutate(prop_ff = round((length_ff / length_river), 2), 
         length_notff = (length_river - length_ff)) |> 
  select(Accessible, prop_ff, length_river, length_ff, length_notff) 
  
```

### Summary by country.
```{r totals-country, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
# Not used. Updated points from Norris et al. 2019.
table_country <- points_bau |> 
  mutate(flag_ff = if_else(Free.flowing == "yes", 1, 0)) |>
  group_by(COUNTRY, Accessible) |> 
  mutate(flag_e = if_else(fem_diff_t42 <= -0.5, 1, 0)) |>
  summarise(length_river = n(), 
            length_end = sum((flag_e)), 
           length_ff = sum(flag_ff), 
           prop_ff = round((sum(flag_ff) / length_river), 2)) |> 
    ungroup() |> 
    mutate(length_not_end = length_river - length_end, 
           prop_end = round((length_end / length_river), 2), 
           length_not_ff = length_river - length_ff
           
) |> 
  select(COUNTRY, Accessible, prop_end, length_river, length_end, length_not_end, 
         prop_ff, length_ff, length_not_ff)

# Sum rows of each column if numeric 
func <- function(z) if (is.numeric(z)) sum(z) else '' 
sumrow <- as.data.frame(lapply(table_country, func))
# Give name to the first element of the new data frame created above
sumrow[1] <- "Total"
# remove unwanted value
sumrow[c(2, 3, 7)] <- NA
```


```{r totals-country-ffr, echo=FALSE, warning=FALSE, message=FALSE}
# All FFR "yes". Need to revise and update.
table_country_ffr <- points_bau_ffr |> 
  mutate(flag_ff = if_else(Free.flowing == "yes", 1, 0)) |>
  group_by(COUNTRY) |> 
  mutate(flag_e = if_else(fem_diff_t42 <= -0.5, 1, 0)) |>
  summarise(pop_change = round(((sum(fem_t42) - sum(fem_t0)) / sum(fem_t0)), 2), 
            change_lcl_95 = Hmisc::smean.cl.boot(fem_diff_t42)["Lower"], 
            change_ucl_95 = Hmisc::smean.cl.boot(fem_diff_t42)["Upper"],
            length_river = n(), 
            length_end = sum((flag_e)), 
           length_ff = sum(flag_ff), 
           prop_ff = round((sum(flag_ff) / length_river), 2)) |> 
    ungroup() |> 
    mutate(length_not_end = length_river - length_end, 
           prop_end = round((length_end / length_river), 2), 
           length_not_ff = length_river - length_ff, 
           change_label = paste(pop_change, " (", round(change_lcl_95, 2), " ", 
                                "\u2013", " ", round(change_ucl_95, 2), ")", sep="")
           
) |> 
  select(COUNTRY, change_label, prop_end, length_river, length_end, length_not_end, 
         prop_ff, length_ff, length_not_ff)

# Sum rows of each column if numeric 
func <- function(z) if (is.numeric(z)) sum(z) else '' 
sumrow_ffr <- as.data.frame(lapply(table_country_ffr, func))
# Give name to the first element of the new data frame created above
sumrow_ffr[1] <- "Total"
# remove unwanted value
sumrow_ffr[c(3, 7)] <- NA
```


```{r make-table-country-ffr, echo=FALSE}
options(knitr.kable.NA = '')

bind_rows(table_country_ffr, sumrow_ffr) |> 
  select(COUNTRY, change_label, prop_end, length_river) |>
knitr::kable(col.names = c("Country", "Population change", 
                           "River length Endangered (prop)", 
                           "River length (km)") 
             )|> 
  kableExtra::column_spec(3:4, width = "3.0cm")
             
```



### Summary by basin.
In the table below "Amazon" includes the Amazon Basin together 
with the Araguaia-Tocantins Basin in Brazil. "Coastal North" covers 
the Guianas and includes the Essequibo Basin. "Coastal South" includes 
part of Maranhão State in Brazil.
```{r totals-basin, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
# Not used. 
# Results from updated Norris et. al. 2019.
table_basin <- points_bau |> 
  group_by(BASIN_N) |> 
  mutate(flag_e = if_else(fem_diff_t42 <= -0.5, 1, 0)) |>
  summarise(pop_change = round(((sum(fem_t42) - sum(fem_t0)) / sum(fem_t0)), 3), 
            length_river = n(), 
            length_end = sum((flag_e))) |> 
    ungroup() |> 
    mutate(length_not_end = length_river - length_end, 
           prop_end = round((length_end / length_river), 2)) |> 
  select(BASIN_N, pop_change, prop_end, length_river, length_end, length_not_end)
```

```{r make-table-basin, echo=FALSE, eval=FALSE}
knitr::kable(table_basin, 
             col.names = c("Basin", "Population change", "Endangered (prop)","river length (km)", 
                           "length End.", "length not End.")) 
             
```


```{r totals-basin-ffr, echo=FALSE, message=FALSE, warning=FALSE}
table_basin_ffr <- points_bau_ffr |> 
  group_by(BASIN_NAME) |> 
  mutate(flag_e = if_else(fem_diff_t42 <= -0.5, 1, 0)) |>
  summarise(pop_start = sum(fem_t0), 
            pop_end = sum(fem_t42), 
            pop_change = round(((sum(fem_t42) - sum(fem_t0)) / sum(fem_t0)), 3),
            change_lcl_95 = Hmisc::smean.cl.boot(fem_diff_t42)["Lower"], 
            change_ucl_95 = Hmisc::smean.cl.boot(fem_diff_t42)["Upper"],
            length_river = n(), 
            length_end = sum((flag_e))) |> 
    ungroup() |> 
    mutate(length_not_end = length_river - length_end, 
           prop_end = round((length_end / length_river), 2), 
           change_label = paste(pop_change, " (", round(change_lcl_95, 2), " ", 
                                "\u2013", " ", round(change_ucl_95, 2), ")", sep="")
           ) |> 
  select(BASIN_NAME, change_label, prop_end, length_river, length_end, length_not_end)
```

```{r make-table-basin-ffr, echo=FALSE}
table_basin_ffr |> 
  select(BASIN_NAME, change_label, prop_end, length_river) |>
knitr::kable(col.names = c("Basin", "population change", 
                           "River length Endangered (prop)", 
                           "River length (km)")) 
             
```


### Summary by major basins and countries.
```{r totals-basin-country, eval = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Not used.
table_basin_country <- points_bau |> 
  group_by(BASIN_N, COUNTRY) |> 
  mutate(flag_e = if_else(fem_diff_t42 <= -0.5, 1, 0)) |>
  summarise(pop_change = round(((sum(fem_t42) - sum(fem_t0)) / sum(fem_t0)), 3), 
            length_river = n(), 
            length_end = sum((flag_e))) |> 
    ungroup() |> 
    mutate(length_not_end = length_river - length_end, 
           prop_end = round((length_end / length_river), 2)) |> 
  select(BASIN_N, COUNTRY, pop_change, prop_end, length_river, length_end, length_not_end) 

```

```{r make-table-basin-country, echo=FALSE, eval=FALSE}
knitr::kable(table_basin_country, 
             col.names = c("Basin", "Country", 
                           "Population change","Endangered (prop)", 
                           "river length (km)", 
                           "length End.", "length not End.")) |> 
  kableExtra::column_spec(3:7, width = "3cm")
             
```


```{r data-basin-country-ffr, echo=FALSE, message=FALSE, warning=FALSE}
# Make summaries
tab_sum <- points_bau_ffr |> 
  dplyr::mutate(flag_EN = if_else(fem_diff_t42 <= -0.5, 1, 0)) |>
  dplyr::group_by(COUNTRY, BASIN_NAME) |> 
  dplyr::summarise(pop_start = sum(fem_t0), 
                   pop_end = sum(fem_t42), 
                   pop_change = round(((sum(fem_t42) - sum(fem_t0)) / sum(fem_t0)), 2),
                   diff_mean = mean(fem_diff_t42), 
                   change_lcl_95 = Hmisc::smean.cl.boot(fem_diff_t42)["Lower"], 
                   change_ucl_95 = Hmisc::smean.cl.boot(fem_diff_t42)["Upper"],
                   diff_median = median(fem_diff_t42),
                   diff_q25 = quantile(fem_diff_t42, probs = 0.25), 
                   diff_q75 = quantile(fem_diff_t42, probs = 0.75),
                   length_river = n(), 
                   length_endangered = sum((flag_EN))) |>
  dplyr::ungroup() |> 
  dplyr::mutate(proportion_endangered = round((length_endangered / length_river), 2), 
                length_label = round((length_river / 1000), 1)
                ) |> 
  dplyr::mutate(threat_status = case_when(diff_mean <= -0.8 ~ "Critically Endangered", 
                                   diff_mean <= -0.5 ~ "Endangered", 
                                   diff_mean <= -0.3 ~ "Vulnerable", 
                                   diff_mean <= -0.2 ~ "Near Threatened"), 
                change_label = paste(pop_change, " (", round(change_lcl_95, 2), " ", 
                                "\u2013", " ", round(change_ucl_95, 2), ")", sep="")
           ) 
#tab_sum$bc <- factor(paste(tab_sum$BASIN_NAME, tab_sum$COUNTRY, sep = "\n"))
# levels(tab_sum$bc)

```


```{r make-table-basin-country-ffr, echo=FALSE}
tab_sum |> 
  filter(length_river > 21) |> 
  droplevels() |>
  select(BASIN_NAME, COUNTRY, change_label, proportion_endangered, 
         length_river) |> 
  arrange(BASIN_NAME, COUNTRY) |>
knitr::kable(col.names = c("Basin", "Country", "population change", 
                           "River length Endangered (prop)", 
                           "river length (km)")) |> 
  kableExtra::column_spec(4:5, width = "3cm")
             
```


The table values can also be presented graphically.
This enables a more comprehensive understanding of the patterns.

```{r fig-basin-country-ffr, echo=FALSE}
# Plot
# Add width factor and fill colour to be consistent across graphs.
mycountry <- sort(unique(tab_sum$COUNTRY))
ncountry <- length(unique(tab_sum$COUNTRY))
palette_country <- grey.colors(n = ncountry)
tab_bc <- tab_sum |> 
  filter(length_river > 21) |> 
  droplevels() |>
  left_join(data.frame(COUNTRY = mycountry, 
                       my_fill = palette_country))
# Now include width factor to be conistent across facets
tab_bc <- tab_bc |> left_join(tab_bc |> 
  group_by(BASIN_NAME) |> summarise(country_count = n()) |> 
  ungroup() |> mutate(width_fac = country_count / max(country_count)) 
)

# check where these errors come from - 
# Orinoco in Brazil, Orinoco in Guyana, Coastal South in French Guiana. 
fig_left <- tab_bc |> 
  filter(BASIN_NAME %in% c("Amazon", "Coastal North")) |>
ggplot(aes(x = COUNTRY, y = pop_change, width = .85*width_fac)) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.2, ymax = -0.3,
            fill="#F6DABF", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.3, ymax = -0.5,
            fill="#DEB688", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.5, ymax = -0.8,
            fill="#C09355", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.8, ymax = -1,
            fill="#A3720E", geom = 'rect', alpha=0.4) + 
  geom_hline(yintercept = 0, colour = "black") +
  geom_col(aes(fill = my_fill)) + 
  geom_errorbar(aes(ymax = change_ucl_95, ymin = change_lcl_95), 
                width = 0.2) + 
  geom_label(aes(x = COUNTRY, y = 0.2, label = length_label)) +
  scale_fill_identity() +  
  facet_wrap(~BASIN_NAME, scales = "free_x", nrow = 2) +
  theme_bw() + 
  labs(y = "Population change", x = "") + 
  #theme(legend.position = "bottom")
  guides(fill = "none")

fig_right <- tab_bc |> 
  filter(BASIN_NAME %in% c("Coastal South", "Orinoco")) |>
ggplot(aes(x = COUNTRY, y = pop_change, width = .85*(width_fac*3))) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.2, ymax = -0.3,
            fill="#F6DABF", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.3, ymax = -0.5,
            fill="#DEB688", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.5, ymax = -0.8,
            fill="#C09355", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.8, ymax = -1,
            fill="#A3720E", geom = 'rect', alpha=0.4) + 
  geom_hline(yintercept = 0, colour = "black") +
  geom_col(aes(fill = my_fill)) + 
  geom_errorbar(aes(ymax = change_ucl_95, ymin = change_lcl_95), 
                width = 0.2) + 
  geom_label(aes(x = COUNTRY, y = 0.2, label = length_label)) + 
  scale_fill_identity() + 
  facet_wrap(~BASIN_NAME, scales = "free_x", nrow = 2) +
  theme_bw() + 
  labs(y = "", x = "") + 
  #theme(legend.position = "bottom")
  guides(fill = "none") 
# put together
fig_left + fig_right + plot_layout(widths = c(3, 1))

```



## Next steps

 - Consultate with researchers to check and validate the mapped results.
