---
title: "Population change map"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interactive-map}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!-- Google Translate widget -->
<div id="google_translate_element" style="float: right;"></div>
<script type="text/javascript">
function googleTranslateElementInit() {
  new google.translate.TranslateElement({pageLanguage: 'en', includedLanguages: 'ar,bn,pt,en,nl,es,fr,de,it,iw,hi,ja,ko,zh-CN'}, 'google_translate_element');
}
</script>
<script type="text/javascript" src="//translate.google.com/translate_a/element.js?cb=googleTranslateElementInit"></script>


```{r setup, echo=FALSE}
library(testmap)
```


```{r load-packages, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(leaflet)
library(leafem)
library(Hmisc)
library(sf)
library(knitr)
library(kableExtra)
library(htmltools) 
library(patchwork)
```
Use the interactive map below to check population changes. Zoom in to see 
where *Podocnemis unilfilis* is Endangered based on IUCN Red List criteria - A3bd.

This is an extension of [Norris et. al. 2019](https://doi.org/10.1016/j.biocon.2019.02.022) that includes:

- Stochastic population projections. 
- Future impacts to populations caused by human acessibility (hunting and habitat change) and actions that reduce river connectivity.

## Map
Due to the number of points, the map can become slow to respond when you zoom in.
Zoom out to a level showing fewer points and you can pan around the map 
to find the area of interest. Then zoom in again to check the coloured points.

```{r make-map-ffr, echo=FALSE, message=FALSE, warning=FALSE}
points_bau4326 <- sf::st_as_sf(points_bau_ffr_map, crs = 3395) |> 
  st_transform(4326) 
levels(points_bau4326$flag_50_42y) <- c("No", "Yes")
points_bau4326_low <- points_bau4326
# label to plot with circle
points_bau4326_low$label <- paste("Prot = ", points_bau4326_low$Protected, 
                                  "Acc = ", points_bau4326_low$Accessible, 
                                  "Free = ", points_bau4326_low$Free.flowing)
# colour palette
leaf_pal <- colorFactor(
  palette = c("#7274C1", "#A3720E"), 
  domain = points_bau4326$flag_50_42y
)

mypal <- c("#7274C1", "#A3720E")

# interactive map. Options added to make panning smoother....
leaflet::leaflet(points_bau4326_low, 
                 options = leafletOptions(preferCanvas = TRUE)) |> 
  addTiles(options = tileOptions(
  updateWhenZooming = FALSE,      # map won't update tiles until zoom is done
  updateWhenIdle = TRUE)) |> 
  addCircles(color = ~leaf_pal(flag_50_42y), 
             popup = ~htmlEscape(label),
             group = "points_bau4326_low"
) |>  
  addCircleMarkers(color = ~leaf_pal(flag_50_42y), 
             stroke = FALSE, fillOpacity = 0.4, 
clusterOptions = markerClusterOptions(), 
group = "points_bau4326"
) |> 
  addLegend("bottomright", pal = leaf_pal, title="Endangered",
            values = ~flag_50_42y,
            group = "en_legend") |> 
  groupOptions("points_bau4326", zoomLevels = 1:7) |> 
  groupOptions("points_bau4326_low", zoomLevels = 8:15) |> 
  groupOptions("en_legend", zoomLevels = 8:15) |> 
  addScaleBar() |> 
  leafem::addMouseCoordinates()
  
```

When you zoom in you will see shaded points. The points follow rivers produced by
remote sensing (Free-Flowing Rivers, 2019: https://doi.org/10.1038/s41586-019-1111-9)
The points are brown where populations are predicted to decline by 50% 
or more within 3 generations (42 years). Brown points therefore represent 
rivers where the species is Endangered, following the IUCN Red List 
population size reduction criteria - A3bd (https://www.iucnredlist.org/about/faqs).

```{r make-map-norris, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
#Map below uses the river points from Norris et. al. 2019.
points_bau4326 <- points_bau |> 
  #filter(BASIN_N == "Orinoco") |>
  arrange(BASIN_N, SUBBASI, BB_ID, REACH_ID) |> 
  filter(row_number() %% 5 == 1) |> st_as_sf(crs = 3395) |> st_transform(4326) 
levels(points_bau4326$flag_50_42y) <- c("No", "Yes")
points_bau4326_low <- points_bau4326
# label to plot with circle
points_bau4326_low$label <- paste("Prot = ", points_bau4326_low$Protected, 
                                  "Acc = ", points_bau4326_low$Accessible, 
                                  "Free = ", points_bau4326_low$Free.flowing)
# colour palette
leaf_pal <- colorFactor(
  palette = c("#7274C1", "#A3720E"), 
  domain = points_bau4326$flag_50_42y
)

mypal <- c("#7274C1", "#A3720E")
 # basemap with tile control. not working.
basemap <- leaflet::leaflet() %>%
  # add different provider tiles
  addProviderTiles(
    "OpenStreetMap",
    # give the layer a name
    group = "OpenStreetMap"
  ) %>%
  addProviderTiles(
    "Stamen.Terrain",
    group = "Stamen.Terrain"
  ) %>%
  addProviderTiles(
    "Esri.WorldImagery",
    group = "Esri.WorldImagery"
  ) %>%
# add a layers control
  addLayersControl(
    baseGroups = c(
      "OpenStreetMap", "Stamen.Terrain", "Esri.WorldImagery"
    ),
    # position it on the topleft
    position = "topleft"
  )

# interactive map. Optins added to make panning smoother....
leaflet::leaflet(points_bau4326_low, 
                 options = leafletOptions(preferCanvas = TRUE)) |> 
  addTiles(options = tileOptions(
  updateWhenZooming = FALSE,      # map won't update tiles until zoom is done
  updateWhenIdle = TRUE)) |> 
  addCircles(color = ~leaf_pal(flag_50_42y), 
             popup = ~htmlEscape(label),
             group = "points_bau4326_low"
) |>  
  addCircleMarkers(color = ~leaf_pal(flag_50_42y), 
             stroke = FALSE, fillOpacity = 0.4, 
clusterOptions = markerClusterOptions(), 
group = "points_bau4326"
) |> 
  addLegend("bottomright", pal = leaf_pal, title="Endangered",
            values = ~flag_50_42y,
            group = "en_legend") |> 
  groupOptions("points_bau4326", zoomLevels = 1:8) |> 
  groupOptions("points_bau4326_low", zoomLevels = 9:15) |> 
  groupOptions("en_legend", zoomLevels = 9:15) |> 
  addScaleBar() |> 
  leafem::addMouseCoordinates()
  
  
```


## Summaries
A summary of the results is presented below. This includes summaries 
by country and major basin.

```{r totals-access, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
# Check length of free flowing by Accessibility.
points_bau |> 
  mutate(flag_ff = if_else(Free.flowing == "yes", 1, 0)) |>
  group_by(Accessible) |> 
  summarise(length_river = n(), 
           length_ff = sum(flag_ff)) |> 
  ungroup() |> 
  mutate(prop_ff = round((length_ff / length_river), 2), 
         length_notff = (length_river - length_ff)) |> 
  select(Accessible, prop_ff, length_river, length_ff, length_notff) 
  
```

### Summary by country.
```{r totals-country, echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
# Not used. Updated points from Norris et al. 2019.
table_country <- points_bau |> 
  mutate(flag_ff = if_else(Free.flowing == "yes", 1, 0)) |>
  group_by(COUNTRY, Accessible) |> 
  mutate(flag_e = if_else(fem_diff_t42 <= -0.5, 1, 0)) |>
  summarise(length_river = n(), 
            length_end = sum((flag_e)), 
           length_ff = sum(flag_ff), 
           prop_ff = round((sum(flag_ff) / length_river), 2)) |> 
    ungroup() |> 
    mutate(length_not_end = length_river - length_end, 
           prop_end = round((length_end / length_river), 2), 
           length_not_ff = length_river - length_ff
           
) |> 
  select(COUNTRY, Accessible, prop_end, length_river, length_end, length_not_end, 
         prop_ff, length_ff, length_not_ff)

# Sum rows of each column if numeric 
func <- function(z) if (is.numeric(z)) sum(z) else '' 
sumrow <- as.data.frame(lapply(table_country, func))
# Give name to the first element of the new data frame created above
sumrow[1] <- "Total"
# remove unwanted value
sumrow[c(2, 3, 7)] <- NA
```
```{r totals-country-ffr, echo=FALSE, warning=FALSE, message=FALSE}
# All "yes". Need to revise and update.
table_country_ffr <- points_bau_ffr |> 
  mutate(flag_ff = if_else(Free.flowing == "yes", 1, 0)) |>
  group_by(COUNTRY) |> 
  mutate(flag_e = if_else(fem_diff_t42 <= -0.5, 1, 0)) |>
  summarise(pop_change = round(((sum(fem_t42) - sum(fem_t0)) / sum(fem_t0)), 3), 
            length_river = n(), 
            length_end = sum((flag_e)), 
           length_ff = sum(flag_ff), 
           prop_ff = round((sum(flag_ff) / length_river), 2)) |> 
    ungroup() |> 
    mutate(length_not_end = length_river - length_end, 
           prop_end = round((length_end / length_river), 2), 
           length_not_ff = length_river - length_ff
           
) |> 
  select(COUNTRY, pop_change, prop_end, length_river, length_end, length_not_end, 
         prop_ff, length_ff, length_not_ff)

# Sum rows of each column if numeric 
func <- function(z) if (is.numeric(z)) sum(z) else '' 
sumrow_ffr <- as.data.frame(lapply(table_country_ffr, func))
# Give name to the first element of the new data frame created above
sumrow_ffr[1] <- "Total"
# remove unwanted value
sumrow_ffr[c(2, 3, 7)] <- NA
```


```{r make-table-country-ffr, echo=FALSE}
options(knitr.kable.NA = '')

bind_rows(table_country_ffr, sumrow_ffr) |> 
  select(COUNTRY, pop_change, prop_end, length_river, length_end, length_not_end) |>
knitr::kable(col.names = c("Country", "Population change", 
                           "Endangered (prop)","river length (km)", 
                           "length End.", "length not End.") 
             )|> 
  kableExtra::column_spec(3:6, width = "2.5cm")
             
```

### Summary by basin.
Using results from updated Norris et. al. 2019.
```{r totals-basin, echo=FALSE, message=FALSE, warning=FALSE}
table_basin <- points_bau |> 
  group_by(BASIN_N) |> 
  mutate(flag_e = if_else(fem_diff_t42 <= -0.5, 1, 0)) |>
  summarise(length_river = n(), 
            length_end = sum((flag_e))) |> 
    ungroup() |> 
    mutate(length_not_end = length_river - length_end, 
           prop_end = round((length_end / length_river), 2)) |> 
  select(BASIN_N, prop_end, length_river, length_end, length_not_end)
```

```{r make-table-basin, echo=FALSE}
knitr::kable(table_basin, 
             col.names = c("Basin", "Endangered (prop)","river length (km)", 
                           "length End.", "length not End.")) 
             
```

Same using Free-flowing rivers data.
```{r totals-basin-ffr, echo=FALSE, message=FALSE, warning=FALSE}
table_basin_ffr <- points_bau_ffr |> 
  group_by(BASIN_NAME) |> 
  mutate(flag_e = if_else(fem_diff_t42 <= -0.5, 1, 0)) |>
  summarise(pop_start = sum(fem_t0), 
            pop_end = sum(fem_t42), 
            pop_change = round(((sum(fem_t42) - sum(fem_t0)) / sum(fem_t0)), 3),
            length_river = n(), 
            length_end = sum((flag_e))) |> 
    ungroup() |> 
    mutate(length_not_end = length_river - length_end, 
           prop_end = round((length_end / length_river), 2)) |> 
  select(BASIN_NAME, pop_change, prop_end, length_river, length_end, length_not_end)
```

```{r make-table-basin-ffr, echo=FALSE}
knitr::kable(table_basin_ffr, 
             col.names = c("Basin", "population change", 
                           "Endangered (prop)","river length (km)", 
                           "length End.", "length not End.")) 
             
```


### Summary by major basins and countries.
```{r totals-basin-country, echo=FALSE, message=FALSE, warning=FALSE}
table_basin_country <- points_bau |> 
  group_by(BASIN_N, COUNTRY) |> 
  mutate(flag_e = if_else(fem_diff_t42 <= -0.5, 1, 0)) |>
  summarise(length_river = n(), 
            length_end = sum((flag_e))) |> 
    ungroup() |> 
    mutate(length_not_end = length_river - length_end, 
           prop_end = round((length_end / length_river), 2)) |> 
  select(BASIN_N, COUNTRY, prop_end, length_river, length_end, length_not_end) 

```

```{r make-table-basin-country, echo=FALSE}
knitr::kable(table_basin_country, 
             col.names = c("Basin", "Country","Endangered (prop)", 
                           "river length (km)", 
                           "length End.", "length not End.")) |> 
  kableExtra::column_spec(3:6, width = "3cm")
             
```

Same with Free-flowing rivers.....plot results
```{r data-basin-country-ffr}
# Make summaries
tab_sum <- points_bau_ffr |> 
  dplyr::mutate(flag_EN = if_else(fem_diff_t42 <= -0.5, 1, 0)) |>
  dplyr::group_by(COUNTRY, BASIN_NAME) |> 
  dplyr::summarise(pop_start = sum(fem_t0), 
                   pop_end = sum(fem_t42), 
                   pop_change = round(((sum(fem_t42) - sum(fem_t0)) / sum(fem_t0)), 3),
                   diff_mean = mean(fem_diff_t42), 
                   change_lcl_95 = Hmisc::smean.cl.boot(fem_diff_t42)["Lower"], 
                   change_ucl_95 = Hmisc::smean.cl.boot(fem_diff_t42)["Upper"],
                   diff_median = median(fem_diff_t42),
                   diff_q25 = quantile(fem_diff_t42, probs = 0.25), 
                   diff_q75 = quantile(fem_diff_t42, probs = 0.75),
                   length_river = n(), 
                   length_endangered = sum((flag_EN))) |>
  dplyr::ungroup() |> 
  dplyr::mutate(proportion_endangered = round((length_endangered / length_river), 2), 
                length_label = round((length_river / 1000), 1)
                ) |> 
  dplyr::mutate(threat_status = case_when(diff_mean <= -0.8 ~ "Critically Endangered", 
                                   diff_mean <= -0.5 ~ "Endangered", 
                                   diff_mean <= -0.3 ~ "Vulnerable", 
                                   diff_mean <= -0.2 ~ "Near Threatened")) 
tab_sum$bc <- factor(paste(tab_sum$BASIN_NAME, tab_sum$COUNTRY, sep = "\n"))
# levels(tab_sum$bc)

```

```{r fig-basin-country-ffr, echo=FALSE}
# Plot
# check where these errors come from - 
# Orinoco in Brazil, Orinoco in Guyana, Coastal South in French Guiana. 
fig_left <- tab_sum |> 
  filter(length_river > 21) |> 
  droplevels() |> 
  filter(BASIN_NAME %in% c("Amazon", "Coastal North")) |>
ggplot(aes(x = COUNTRY, y = pop_change)) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.2, ymax = -0.3,
            fill="#F6DABF", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.3, ymax = -0.5,
            fill="#DEB688", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.5, ymax = -0.8,
            fill="#C09355", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.8, ymax = -1,
            fill="#A3720E", geom = 'rect', alpha=0.4) + 
  geom_hline(yintercept = 0, colour = "black") +
  geom_col(aes(fill = COUNTRY)) + 
  geom_errorbar(aes(ymax = change_ucl_95, ymin = change_lcl_95), 
                width = 0.2) + 
  geom_label(aes(x = COUNTRY, y = 0.2, label = length_label)) +
  scale_fill_grey() + 
  facet_wrap(~BASIN_NAME, scales = "free_x", nrow = 2) +
  theme_bw() + 
  labs(y = "Population change", x = "") + 
  #theme(legend.position = "bottom")
  guides(fill = "none")
fig_right <- tab_sum |> 
  filter(length_river > 21) |> 
  droplevels() |> 
  filter(BASIN_NAME %in% c("Coastal South", "Orinoco")) |>
ggplot(aes(x = COUNTRY, y = pop_change)) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.2, ymax = -0.3,
            fill="#F6DABF", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.3, ymax = -0.5,
            fill="#DEB688", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.5, ymax = -0.8,
            fill="#C09355", geom = 'rect', alpha=0.4) + 
  annotate(xmin=-Inf, xmax=Inf, ymin = -0.8, ymax = -1,
            fill="#A3720E", geom = 'rect', alpha=0.4) + 
  geom_hline(yintercept = 0, colour = "black") +
  geom_col(aes(fill = COUNTRY)) + 
  geom_errorbar(aes(ymax = change_ucl_95, ymin = change_lcl_95), 
                width = 0.2) + 
  geom_label(aes(x = COUNTRY, y = 0.2, label = length_label)) +
  scale_fill_grey() + 
  facet_wrap(~BASIN_NAME, scales = "free_x", nrow = 2) +
  theme_bw() + 
  labs(y = "", x = "") + 
  #theme(legend.position = "bottom")
  guides(fill = "none") 
# put together
fig_left + fig_right + plot_layout(widths = c(3, 1))

```

Table showing data values used in graph.

```{r make-table-basin-country-ffr, echo=FALSE}
tab_sum |> 
  filter(length_river > 21) |> 
  droplevels() |>
  select(BASIN_NAME, COUNTRY, pop_change, length_river) |> 
  arrange(BASIN_NAME, COUNTRY) |>
knitr::kable(col.names = c("Basin", "Country", "population change", 
                           "river length (km)")) |> 
  kableExtra::column_spec(3:4, width = "3cm")
             
```


## Next steps

 - Update using points from Free Flowing Rivers    
 <https://www.worldwildlife.org/pages/free-flowing-rivers>
