---
title: "Population change map"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interactive-map}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(testmap)
```


```{r load-packages, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(leaflet)
library(leafem)
library(sf)
```
Use the interactive map below to check population changes. Zoom in to see 
where *Podocnemis unilfilis* is Endangered based on IUCN Redlist criteria - A3bd.

This is an extension of [Norris et al 2019](https://doi.org/10.1016/j.biocon.2019.02.022) that includes:

- stochastic population projections 
- threats caused by human acessibility (hunting and habitat change) and actions that reduce river connectivity.

Due to the number of points, the map can become slow to respond when you zoom in.
Zoom out to a level showing fewer points and you can pan around the map 
to find the area of interest. Then zoom in again to check the coloured points.
```{r make-map, echo=FALSE, message=FALSE, warning=FALSE}
points_bau4326 <- points_bau |> 
  #filter(BASIN_N == "Orinoco") |>
  arrange(BASIN_N, SUBBASI, BB_ID, REACH_ID) |> 
  filter(row_number() %% 5 == 1) |> st_as_sf() |> st_transform(4326) 
levels(points_bau4326$flag_50_42y) <- c("No", "Yes")
points_bau4326_low <- points_bau4326
# colour palette
leaf_pal <- colorFactor(
  palette = c("#7274C1", "#A3720E"), 
  domain = points_bau4326$flag_50_42y
)

mypal <- c("#7274C1", "#A3720E")
 # basemap with tile control. not working.
basemap <- leaflet::leaflet() %>%
  # add different provider tiles
  addProviderTiles(
    "OpenStreetMap",
    # give the layer a name
    group = "OpenStreetMap"
  ) %>%
  addProviderTiles(
    "Stamen.Terrain",
    group = "Stamen.Terrain"
  ) %>%
  addProviderTiles(
    "Esri.WorldImagery",
    group = "Esri.WorldImagery"
  ) %>%
# add a layers control
  addLayersControl(
    baseGroups = c(
      "OpenStreetMap", "Stamen.Terrain", "Esri.WorldImagery"
    ),
    # position it on the topleft
    position = "topleft"
  )

# interactive map
leaflet::leaflet(points_bau4326_low) |> 
  addTiles() |> 
  addCircles(color = ~leaf_pal(flag_50_42y), 
             group = "points_bau4326_low"
) |>  
  addCircleMarkers(color = ~leaf_pal(flag_50_42y), 
             stroke = FALSE, fillOpacity = 0.4, 
clusterOptions = markerClusterOptions(), 
group = "points_bau4326"
) |> 
  addLegend("bottomright", pal = leaf_pal, title="Endangered",
            values = ~flag_50_42y,
            group = "en_legend") |> 
  groupOptions("points_bau4326", zoomLevels = 1:8) |> 
  groupOptions("points_bau4326_low", zoomLevels = 9:15) |> 
  groupOptions("en_legend", zoomLevels = 9:15) |> 
  addScaleBar() |> 
  leafem::addMouseCoordinates()
  
  
```

When you zoom in you will see shaded points. Points are brown where 
populations are expected to decline by 50% or more within 3 generations (42 years).
Brown points therefore represent rivers where the species is Endangered, 
following the IUCN Redlist criteria - A3bd.
